# Reproducible Research: Peer Assessment 1
written by Unshar, June 2014


## Loading and preprocessing the data
```{r loaddata, echo=TRUE}
# Install function for required packages    
packages<-function(x){
  x<-as.character(match.call()[[2]])
  if (!require(x,character.only=TRUE)){
    install.packages(pkgs=x,repos="http://cran.r-project.org")
    require(x,character.only=TRUE)
  }
}
packages(ggplot2)
packages(scales)
packages(lattice)

#Load data
  dir<-getwd()
  path<-paste(c(dir,"/activity.csv"), collapse='')
  mydata<-read.csv(path, na.string="NA")
```


## What is mean total number of steps taken per day?
*Missing values are being ignored as directed by the assignment instructions.

```{r q1Mean, echo=TRUE}
#Aggregate steps by date and calculate mean and median
StepsTotal <-aggregate(steps ~ date, data = mydata, sum, na.rm = T)  
colnames(StepsTotal) <- c("Date", "Total")
  x<-mean(StepsTotal$Total)
  y<-median(StepsTotal$Total)
```
### The mean for the total number of steps for all days is `r x`.
### The median for the total number of steps for all days is `r y`.

#### Mean breakdown by day*
```{r q1Meanbrk, echo=TRUE}
StepsMean <-aggregate(steps ~ date, data = mydata, mean, na.rm = T) 
  colnames(StepsMean) <- c("Date", "Mean")
  StepsMean
```

#### Median breakdown by day*
```{r q1Meadianbrk, echo=TRUE}
StepsMedian <-aggregate(steps ~ date, data = mydata, median, na.rm = T)
  colnames(StepsMedian) <- c("Date", "Median")
  StepsMedian
```

### Histogram*
```{r q1Hist, echo=TRUE}
#Aggregate steps by date
  StepsTotal$names <- as.Date(StepsTotal$Date, format="%Y-%m-%d")

#Histogram total steps bins
  m <- ggplot(StepsTotal, aes(x=Total))
    m + geom_histogram(binwidth = 1000, aes(fill = ..count..)) +
    scale_fill_gradient("Days", low = "red", high = "green")+
    ylab("Days(Count)") + xlab("Total steps")
```

## What is the average daily activity pattern?
```{r q2, echo=TRUE}
IntervalsMean <-aggregate(steps ~ interval, data = mydata, mean, na.rm = T)
  with(IntervalsMean, plot(interval,
                         steps, 
                         type = "l",
                         xlab = "Interval",
                         ylab = "Average number of steps"))   

```  
### Which 5-min interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r q2answer, echo=TRUE}

n <- length(IntervalsMean$steps)
  y<-sort(IntervalsMean$steps,partial=n-0)[n-0]
    x<-subset(IntervalsMean,IntervalsMean$steps==y)
      result<-x[1,1]
```
###  ANSWER: The `r result`th interval contains the maximum number of steps (`r y`) on average across all the days in this dataset.


## Imputing missing values
```{r q4NAs, echo=TRUE}
#Number os NAs
  x<-sum( is.na(mydata$steps) )
```
### The total number of missing values in the dataset is: `r x`
According to this assignments instructions the strategy for filling in NAs does not need to be sophisticated. Since 63% of the observations are zero (against 24% non-NAs and non-zeros ad 13% NAs), the strategy selected was to replace NAs with the same value as majority of observations.

```{r q4Replace NAs, echo=TRUE}
#Replace NAs with zeros
  mydata$steps[is.na(mydata$steps)] <- 0 

##Aggregate steps by date and calculate mean and median
  StepsTotal <-aggregate(steps ~ date, data = mydata, sum, na.rm = T)  
colnames(StepsTotal) <- c("Date", "Total")
  x<-mean(StepsTotal$Total)
  y<-median(StepsTotal$Total)
```

### Histogram of the total number of steps taken each day
```{r q4Hist, echo=TRUE}
#Histogram total steps bins
  m <- ggplot(StepsTotal, aes(x=Total))
    m + geom_histogram(binwidth = 1000, aes(fill = ..count..)) +
    scale_fill_gradient("Days", low = "red", high = "green")+
    ylab("Days(Count)") + xlab("Total steps")
```

#### Updated Mean breakdown by day
```{r q4Meanbrk, echo=TRUE}
StepsMean <-aggregate(steps ~ date, data = mydata, mean, na.rm = T) 
  colnames(StepsMean) <- c("Date", "Mean")
  StepsMean
```

#### Updated Median breakdown by day
```{r q4Meadianbrk, echo=TRUE}
StepsMedian <-aggregate(steps ~ date, data = mydata, median, na.rm = T)
  colnames(StepsMedian) <- c("Date", "Median")
  StepsMedian
```


## Are there differences in activity patterns between weekdays and weekends?
```{r q5, echo=TRUE}
#Create Weekday field
  mydata$day<-weekdays(as.Date(mydata$date))
  mydata$flagday<- ifelse(mydata$day=="Saturday"|mydata$day=="Sunday", "Weekend","Weekday")

#Aggregate steps by Weekday and Interval
  IntervalsMeanDay <-aggregate(mydata$steps, list(day=mydata$flagday, intervals=mydata$interval), mean, na.rm=TRUE)

#Calculate the total mean of the two different sets (Weekend and Weekdays)
x<-(subset(IntervalsMeanDay,IntervalsMeanDay$day=="Weekend"))
xstepswkd<-mean(x$x)
x<-(subset(IntervalsMeanDay,IntervalsMeanDay$day=="Weekday"))
xstepsweek<-mean(x$x)
    
#plot by weekday
  dayplot<- xyplot(x~intervals |day, data=IntervalsMeanDay, 
                 ylim=range(IntervalsMeanDay$x),
                 main="Average steps by intervals", 
                 ylab="Average steps",  
                 xlab="Intervals", type="l",
                 par.strip.text=list(cex=0.7), 
                 par.settings=list(axis.text=list(cex=0.7)))
  print(dayplot)
```  

### Comment: Subject appears to move more constanly over weekends, taking an average of `r xstepswkd` steps vs `r xstepsweek` steps during the week.